---
title: "R Notebook"
output: html_notebook
---

# Packages
```{r}
require(tidyverse)
require(terra)
require(tidyterra)
require(sf)
require(adehabitatHR)
require(adehabitatLT)
require(landscapemetrics)
require(reshape2)

```



# Challenge 1 (4 points)

Create a table with 6 rows and 5 columns. Each row should represent one panther. The first column should represent the panther's ID, and the other 4 columns should represent the estimated 95% home range. Make sure that all of your units for your home range estimates are the same. Ensure that the final table outputs from the code below.

```{r}
# Data

panthers = st_read('/vsicurl/https://github.com/ValenteJJ/SpatialEcology/raw/main/Week10/panthers.shp') %>% 
  mutate(CatID = as.factor(CatID))

summary(panthers)
CatID = unique(panthers$CatID)
CatID

panthersSp = as(panthers, 'Spatial')

# LULC

land = rast('https://github.com/ValenteJJ/SpatialEcology/raw/main/Week10/panther_landcover.tif')

classification = read.table('https://raw.githubusercontent.com/ValenteJJ/SpatialEcology/main/Week10/landcover%20reclass.txt', header=T) 

head(classification)

unique(classification$Description2)

land = classify(land, classification[,c(1,3)])
land = categories(land, value=unique(classification[,c(3,4)]))
plot(land)

wetForest = land
values(wetForest) = 0
wetForest[land %in% c(10,12)] = 1


probMatrix = focalMat(wetForest, 5000, type='circle', fillNA=FALSE)
wetFocal = focal(wetForest, probMatrix, fun='sum', na.rm=T)

dryForest = land
values(dryForest) = 0
dryForest[land %in% c(11, 13)] = 1

probMatrix = focalMat(dryForest, 5000, type='circle', fillNA=FALSE)
dryFocal = focal(dryForest, probMatrix, fun='sum', na.rm=T)

layers = c(land, wetFocal, dryFocal)
names(layers) = c('landcover', 'wetForest', 'dryForest')
plot(layers)

```

```{r}
# MCP

mcp95 = mcp(panthersSp[,'CatID'], percent = 95, unin='m', unout='m2')
mcp95Sf = st_as_sf(mcp95)
mcp95

# Kdens

kernelHrefEpa = kernelUD(panthersSp[,'CatID'], h='href', kern='epa')
image(kernelHrefEpa)

## Kdens Bandwidth
kernelHrefEpa[[2]]@h
## Kdens plot of 2nd panther?
plot(kernelHrefEpa[[2]])
## Kdens the rest...
kernel.area(kernelHrefEpa)

plot(getverticeshr(kernelHrefEpa))

kde95Sf = st_as_sf(getverticeshr(kernelHrefEpa, percent=95))
kde95Sf

# Brown

substrRight = function(x, n){
  substr(x, nchar(x) - n+1, nchar(x))
}

panthersSp = panthers %>% 
  mutate(Juldate = as.character(Juldate)) %>% 
  mutate(date = as.numeric(substrRight(Juldate, 3))) %>% 
  mutate(Date = as.Date(date, origin=as.Date("2006-01-01"))) %>% 
  mutate(Date = as.POSIXct(Date, "%Y-%m-%d")) %>% 
  as('Spatial')

pantherLtraj = as.ltraj(xy=coordinates(panthersSp), date=panthersSp$Date, id=panthersSp$CatID, typeII=T)

plot(pantherLtraj)

sigma1 = liker(pantherLtraj, sig2=450, rangesig1=c(2, 100))

sigma1

# 100

bb100 = kernelbb(pantherLtraj[1], sig=13.8, sig2=450, grid=500)
plot(bb100)

bb95Sf100 = st_as_sf(getverticeshr(bb100, percent=95))

# 130

bb130 = kernelbb(pantherLtraj[2], sig=18.1, sig2=450, grid=500)
plot(bb130)

bb95Sf130 = st_as_sf(getverticeshr(bb130, percent=95))

# 131

bb131 = kernelbb(pantherLtraj[3], sig=13.4, sig2=450, grid=500)
plot(bb131)

bb95Sf131 = st_as_sf(getverticeshr(bb131, percent=95))

# 137

bb137 = kernelbb(pantherLtraj[4], sig=15.1, sig2=450, grid=500)
plot(bb137)

bb95Sf137 = st_as_sf(getverticeshr(bb137, percent=95))

# 143

bb143 = kernelbb(pantherLtraj[5], sig=10.8, sig2=450, grid=500)
plot(bb143)

bb95Sf143 = st_as_sf(getverticeshr(bb143, percent=95))

# 147

bb147 = kernelbb(pantherLtraj[6], sig=7.2, sig2=450, grid=500)
plot(bb147)

bb95Sf147 = st_as_sf(getverticeshr(bb147, percent=95))

# bind

brown = c(bb95Sf100$area, bb95Sf130$area, bb95Sf131$area, bb95Sf137$area, bb95Sf143$area, bb95Sf147$area)
```

```{r}
# Locoh 147

panther147 = panthersSp[panthersSp$CatID==147,]

kInit = round(nrow(coordinates(panther147))^0.5, 0)
aInit = round(max(dist(coordinates(panther147))),0)

kSearch = seq(kInit, 10*kInit, by=5) #number of points
aSearch = seq(aInit, 2*aInit, by=3000) #distance in m
kSearch
aSearch

locohArange = LoCoH.a.area(SpatialPoints(coordinates(panther147)), arange=aSearch)
locohKrange = LoCoH.k.area(SpatialPoints(coordinates(panther147)), krange=kSearch)

aSearch[5]
kSearch[11]

locohA = LoCoH.a(SpatialPoints(coordinates(panther147)), a=aSearch[5])
plot(locohA)

locohK = LoCoH.k(SpatialPoints(coordinates(panther147)), k = kSearch[11])
plot(locohK)

locohASf = st_as_sf(locohA)
locohKSf147 = st_as_sf(locohK)

locohA92Sf = locohASf[97,]
locohK95Sf147 = locohKSf147[83,]

# Locoh 100

panther100 = panthersSp[panthersSp$CatID==100,]

kInit100 = round(nrow(coordinates(panther100))^0.5, 0)

kSearch100 = seq(kInit100, 10*kInit100, by=5) #number of points
kSearch100

locohKrange100 = LoCoH.k.area(SpatialPoints(coordinates(panther100)), krange=kSearch100)

kSearch100[12]

locohK100 = LoCoH.k(SpatialPoints(coordinates(panther100)), k = kSearch100[12])
plot(locohK100)

locohKSf100 = st_as_sf(locohK100)

locohK95Sf100 = locohKSf100[17,]

# Locoh 130

panther130 = panthersSp[panthersSp$CatID==130,]

kInit130 = round(nrow(coordinates(panther130))^0.45, 0)

kSearch130 = seq(kInit130, 10*kInit130, by=5) #number of points
kSearch130

locohKrange130 = LoCoH.k.area(SpatialPoints(coordinates(panther130)), krange=kSearch130)

kSearch130[3]

locohK130 = LoCoH.k(SpatialPoints(coordinates(panther130)), k = kSearch130[3])
plot(locohK130)

locohKSf130 = st_as_sf(locohK130)

locohK95Sf130 = locohKSf130[80,]

# Locoh 131

panther131 = panthersSp[panthersSp$CatID==131,]

kInit131 = round(nrow(coordinates(panther131))^0.5, 0)

kSearch131 = seq(kInit131, 10*kInit131, by=5) #number of points
kSearch131

locohKrange131 = LoCoH.k.area(SpatialPoints(coordinates(panther131)), krange=kSearch131)

kSearch131[5]

locohK131 = LoCoH.k(SpatialPoints(coordinates(panther131)), k = kSearch131[5])
plot(locohK131)

locohKSf131 = st_as_sf(locohK131)

locohK95Sf131 = locohKSf131[64,]

# Locoh 137

panther137 = panthersSp[panthersSp$CatID==137,]

kInit137 = round(nrow(coordinates(panther137))^0.5, 0)

kSearch137 = seq(kInit137, 10*kInit137, by=5) #number of points
kSearch137

locohKrange137 = LoCoH.k.area(SpatialPoints(coordinates(panther137)), krange=kSearch137)

kSearch137[2]

locohK137 = LoCoH.k(SpatialPoints(coordinates(panther137)), k = kSearch137[2])
plot(locohK137)

locohKSf137 = st_as_sf(locohK137)

locohK95Sf137 = locohKSf137[103,]

# Locoh 143

panther143 = panthersSp[panthersSp$CatID==143,]

kInit143 = round(nrow(coordinates(panther143))^0.5, 0)

kSearch143 = seq(kInit143, 10*kInit143, by=5) #number of points
kSearch143

locohKrange143 = LoCoH.k.area(SpatialPoints(coordinates(panther143)), krange=kSearch143)

kSearch143[10]

locohK143 = LoCoH.k(SpatialPoints(coordinates(panther143)), k = kSearch143[10])
plot(locohK143)

locohKSf143 = st_as_sf(locohK143)

locohK95Sf143 = locohKSf143[92,]

Locoh = c(11479.33, 85078.74, 19599.4, 17185.75, 49630.19, 116343.9)
Locoh
```


```{r}



# Table
ModelsTable = data.frame(matrix(nrow=6, ncol=5))
colnames(ModelsTable)[1] <- "CatID"
colnames(ModelsTable)[2] <- "MCP"
colnames(ModelsTable)[3] <- "Kdens"
colnames(ModelsTable)[4] <- "Locoh"
colnames(ModelsTable)[5] <- "Brown"

ModelsTable$CatID = c(100, 130, 131, 137, 143, 147)
ModelsTable$MCP = mcp95$area
ModelsTable$Kdens = kde95Sf$area
ModelsTable$Locoh = Locoh
ModelsTable$Brown = brown

```

No text necessary.


# Challenge 2 (4 points)

Build a boxplot showing the distribution of home range sizes. Your x-axis should have 4 categories (one for each home range type), and your y-axis should show the distribution of the home ranges for the 6 panthers. Which home range estimates tend to be largest? Which ones tend to be smallest? What is your intuitive explanation for why this pattern occurs?

```{r}
# Code here
MCPcats = ModelsTable$MCP
Kdenscats = ModelsTable$Kdens
Locohcats = ModelsTable$Locoh
Browncats = ModelsTable$Brown


boxplot(MCPcats, Kdenscats, Locohcats, Browncats, log="y",
        names = c("MCP", "Kdens", "Locoh", "Brown"))


```

MCP had much larger HR values than other methods, and so for visual purposes, the y axis (area) was log transformed. Much larger area values in MCP was th result of it's broad strokes approach, where 95% of the core area between points is included. The lower values in other methods reflect the more precision based approach of defining that core area relative to movement, or arrangement of point location data. Locoh seems to the most even distributed range of values for all cats, whereas Brown has the largest. I'm assuming that's whats going on with the dot over it, but its max value (cat 130) is several degrees of magnitude larger than all other cats with this method. Because it homes in on the temporal aspect more, I think cat 130 spent enough time in the long path it took outside of the core area, that this was considered a part of it's home range whereas others gave it less weight. Other than this outlier, it's range is the smallest. If time/distance between points was relatively consistant for all cats, I could see this sort of 'standardizing' the dataset to a smaller range. It makes sense that Locoh would have lower values in general, since it's whole thing is to create the smallest possible polygons for contiguity. I think more variance in area is captured by kernel density through bandwidth functions, etc., and brownian models, so that it's reasonable to attribute more area to the amount of variance in those methods. Probably a huge oversimplification, but I'm drinking from a fire hydrant on how exactly this works in the land of statistics.      

# Challenge 3 (3 points)

Choose 1 panther and generate a 100% home range based on the MCP. Calculate (and report) the proportion of this polygon comprised of wet and dry forest. Do you think these amounts are representative of the wet and dry forest available to your panther? Why or why not?

```{r}
# MCP

mcp100_137 = mcp(panther137, percent = 100, unin='m', unout='m2')
mcp100Sf_137 = st_as_sf(mcp100_137)
mcp100_137
mcp100Sf_137
plot(mcp100Sf_137)
plot(land)

mcp100Sf_137_wf = extract(wetForest, mcp100Sf_137)
mcp100Sf_137_wf
sum(mcp100Sf_137_wf$Description2)

mcp100Sf_137_df = extract(dryForest, mcp100Sf_137)
mcp100Sf_137_df
sum(mcp100Sf_137_df$Description2)






```


Wet Forest = 54.43218 /1143 -> 5.07%
Dry Forest = 162.265 /1143 -> 14.61%

No. Being a 100% home range as a MCP, there is no doubt that it doesn't utilize all the space within this area. There might be forested areas it doesn't utilize at all depending on whats around them and we might get a better idea with a different home range model.


# Challenge 4 (6 points)

Using the same panther you chose for Challenge 3, generate 10 sf objects, one representing each of the 10%, 20%, 30%, ..., 90%, and 100% home ranges based on a Brownian bridge model. Extract the proportion of each of those territories comprised of dry forest and wet forest. Now generate two line graphs, one for wet forest and one for dry. On the x-axis, plot increasing home range size (10-100%). On the y-axis plot the proportion of the territory comprised of wet/dry forest, and link these with a line. Finally, add a horizontal line that shows the rough proportion of that forest type available to the panther (based on challenge 3). Show your output below.

```{r}
bb10Sf137 = st_as_sf(getverticeshr(bb137, percent=10))
bb20Sf137 = st_as_sf(getverticeshr(bb137, percent=20))
bb30Sf137 = st_as_sf(getverticeshr(bb137, percent=30))
bb40Sf137 = st_as_sf(getverticeshr(bb137, percent=40))
bb50Sf137 = st_as_sf(getverticeshr(bb137, percent=50))
bb60Sf137 = st_as_sf(getverticeshr(bb137, percent=60))
bb70Sf137 = st_as_sf(getverticeshr(bb137, percent=70))
bb80Sf137 = st_as_sf(getverticeshr(bb137, percent=80))
bb90Sf137 = st_as_sf(getverticeshr(bb137, percent=90))
bb100Sf137 = st_as_sf(getverticeshr(bb137, percent=99))

#10
bb10Sf137_wf = extract(wetForest, bb10Sf137)
bb10Sf137_wf
sum(bb10Sf137_wf$Description2)
bb10Sf137_df = extract(dryForest, bb10Sf137)
bb10Sf137_df
sum(bb10Sf137_df$Description2)
#20
bb20Sf137_wf = extract(wetForest, bb20Sf137)
bb20Sf137_wf
sum(bb20Sf137_wf$Description2)
bb20Sf137_df = extract(dryForest, bb20Sf137)
bb20Sf137_df
sum(bb20Sf137_df$Description2)
#30
bb30Sf137_wf = extract(wetForest, bb30Sf137)
bb30Sf137_wf
sum(bb30Sf137_wf$Description2)
bb30Sf137_df = extract(dryForest, bb30Sf137)
bb30Sf137_df
sum(bb30Sf137_df$Description2)
#40
bb40Sf137_wf = extract(wetForest, bb40Sf137)
bb40Sf137_wf
sum(bb40Sf137_wf$Description2)
bb40Sf137_df = extract(dryForest, bb40Sf137)
bb40Sf137_df
sum(bb40Sf137_df$Description2)
#50
bb50Sf137_wf = extract(wetForest, bb50Sf137)
bb50Sf137_wf
sum(bb50Sf137_wf$Description2)
bb50Sf137_df = extract(dryForest, bb50Sf137)
bb50Sf137_df
sum(bb50Sf137_df$Description2)
#60
bb60Sf137_wf = extract(wetForest, bb60Sf137)
bb60Sf137_wf
sum(bb60Sf137_wf$Description2)
bb60Sf137_df = extract(dryForest, bb60Sf137)
bb60Sf137_df
sum(bb60Sf137_df$Description2)
#70
bb70Sf137_wf = extract(wetForest, bb70Sf137)
bb70Sf137_wf
sum(bb70Sf137_wf$Description2)
bb70Sf137_df = extract(dryForest, bb70Sf137)
bb70Sf137_df
sum(bb70Sf137_df$Description2)
#80
bb80Sf137_wf = extract(wetForest, bb80Sf137)
bb80Sf137_wf
sum(bb80Sf137_wf$Description2)
bb80Sf137_df = extract(dryForest, bb80Sf137)
bb80Sf137_df
sum(bb80Sf137_df$Description2)
#90
bb90Sf137_wf = extract(wetForest, bb90Sf137)
bb90Sf137_wf
sum(bb90Sf137_wf$Description2)
bb90Sf137_df = extract(dryForest, bb90Sf137)
bb90Sf137_df
sum(bb90Sf137_df$Description2)
#100
bb100Sf137_wf = extract(wetForest, bb100Sf137)
bb100Sf137_wf
sum(bb100Sf137_wf$Description2)
bb100Sf137_df = extract(dryForest, bb100Sf137)
bb100Sf137_df
sum(bb100Sf137_df$Description2)

#Table
BB_Table = data.frame(matrix(nrow=10, ncol=3))
colnames(BB_Table)[1] <- "ID"
colnames(BB_Table)[2] <- "WF"
colnames(BB_Table)[3] <- "DF"

BB_Table$ID = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100)
BB_Table$WF = c(0.14, 0.11, 0.079, 0.069, 0.058, 0.051, 0.045, 0.044, .049, 0.049)
BB_Table$DF = c(0.20, 0.20, 0.23, .21, 0.19, 0.17, 0.15, 0.14, 0.14, 0.11)

plot(BB_Table$ID, BB_Table$WF)
plot(BB_Table$ID, BB_Table$DF)

BB_Table.long <- melt(BB_Table, id = "ID", measure = c("WF", "DF"))
ggplot(BB_Table.long, aes(ID, value, colour = variable)) + geom_line() + 
  geom_hline(yintercept = 0.0507) + geom_hline(yintercept = 0.146)


```



# Challenge 5 (3 points)

It seems that the estimates of a 100% MCP  considered non-forested areas outside of core areas too much. That's why I take it that there's more of either forest type available in lesser home range areas for Brownian models. My take is that both start declining in availability to some extent as home ranges exceed a 35% core area. I would expect the home range to have more wet forest available if the panther truly liked it better, but I know that's too simple. I think the most important thing to note is how dry forest increases with core area initially, and how wet forest declines. They both decline after 35%, but it seems like its core area is more dry forest dominant until there's no more to go around.There's still not enough info to answer it trully, but the evidence here supports dry forest. And yes. I am not surprised that this amount wasn't truly available to them at 100%, just because of the nature of MCP. 


