---
title: "Lab 4 Assignment - Landcover"
output:
  html_document:
    df_print: paged
---



```{r}
rm(list=ls())
require(Voss)
require(tidyverse)
require(terra)
require(FedData)
require(sf)
require(tidyterra)
require(landscapemetrics)
```

## Challenge 1 (4 points)

**The landscapemetrics package has functions for calculating 12 patch-level metrics. Calculate all 12 of these for every forest patch in our nlcdSimple raster (using the 8-direction rule). This will result in 12 unique values associated with each forest patch. Use the chart.Correlation() function from the PerformanceAnalytics package to examine the correlation among all 12 of these variables. What patterns do you notice, and what do they tell you about the uniqueness or redundancy of the patch-level metrics you calculated?**
### Setup
```{r}
studyArea = st_read('/vsicurl/https://github.com/ValenteJJ/SpatialEcology/raw/main/Week4/studyArea.shp')
nlcd = get_nlcd(studyArea, label='AlLandscape', year=2011)

# values(nlcd) = as.character(values(nlcd))

plot(nlcd)

nlcdSimple = nlcd
nlcdSimple[nlcdSimple==11] = 1 #Wet areas are a 1 now
nlcdSimple[nlcdSimple %in% c(21, 22, 23, 24)] = 2 #All developed areas are 2
nlcdSimple[nlcdSimple %in% c(31, 52)] = 3 #Barren land and shrub/scrub are 3
nlcdSimple[nlcdSimple %in% c(41,42,43)] = 4 #All forest types are 4
nlcdSimple[nlcdSimple == 71] = 5 #Grassland is 5
nlcdSimple[nlcdSimple %in% c(81,82)] = 6 #And agriculture is 6

#Give these numbers category names
tmp = data.frame('ID' = c(1, 2, 3, 4, 5, 6),
                 'category' = c('wetland', 'developed', 'open', 'forest', 'grassland', 'agriculture'))
nlcdSimple = categories(nlcdSimple, value=tmp)

#And plot the new raster
ggplot(nlcdSimple, aes(x=x, y=y, fill=category)) +
  geom_raster()+
  scale_fill_manual(values=c('blue', 'black', 'gray', 'darkolivegreen', 'orange', 'yellow'))

res(nlcdSimple)
ext(nlcdSimple)
levels(nlcdSimple)


forest = nlcdSimple %>% 
  setValues(NA)
  
forest[nlcdSimple ==4] = 1

plot(forest)

forestPatchId = patches(forest, directions=8, zeroAsNA=T, allowGaps=F)

plot(forestPatchId)
```

### Code

```{r}
# Patch Area
## literally just the area of all patches
patchArea = lsm_p_area(forest, directions=8)
patchArea

# Core Area Index
## the percentage of a patch that is core area. describes patch area and shape simultaneous. relative metric, independent of total area.
patchCAI = lsm_p_cai(forest, directions = 8)
patchCAI

# Core Area
## the area within a patch that is not on the edge...
patchCore = lsm_p_core(forest, directions=8)
patchCore

# Patch Circle
## ratio between the patch area and the smallest circumscribing circle of the patch. measure of compactness of the patch.
patchCircle = lsm_p_circle(forest, directions = 8)
patchCircle

# Contiguity
## weighted average which favors rook over queen for connectivity index. 
patchCont = lsm_p_contig(forest, directions = 8)
patchCont

# Euclidean Nearest-Neighbor Distance
## the distance to the nearest neighboring patch of the same class. edge to edge. simple isolation metrics.
patchEnn = lsm_p_enn(forest, directions=8)
patchEnn

# Fractal dimension index
## Shape metric, index based on the patch perimeter and the patch area and describes the patch complexity. Fractals are weird shapes you see when you listen to TOOL...not circles and squares.
patchFDI = lsm_p_frac(forest, directions = 8)
patchFDI

# Radius of Gyration (Area and edge metric)
## Area and edge metric. The distance from each cell to the patch centroid is based on cell center to centroid distances. Characterizes both the patch area and compactness.
patchGyrate = lsm_p_gyrate(forest, directions = 8)
patchGyrate

# number of core areas
## Core area metric. Describes patch area and shape simultaneously. Patches within patch. a compact shape will contain less disjunct core areas and vice versa.
patchNcore = lsm_p_ncore(forest, directions = 8)
patchNcore

# perimeter area ratio
## self evident. Scale dependent because of variability of shapes with perimeter.
patchPara = lsm_p_para(forest, directions=8)
patchPara

# Perimeter
## Just the perimeter. Don't discount how shape influences on top of area.
patchPerim = lsm_p_perim(forest, directions=8)
patchPerim

# Shape Index
## Shape complexity
patchShape = lsm_p_shape(forest, directions = 8)
patchShape

tmp = rbind(patchShape, patchPerim, patchPara) %>% 
  pivot_wider(names_from=metric, values_from=value) %>% 
  select(shape:para)

PerformanceAnalytics::chart.Correlation(tmp, histogram=F)

# Chart correlation matrix
#This is where we are going to store the output values
IDK how to get these into a dataframe. %>% 
  as.data.frame() %>% 
  select(patchArea, patchCAI, patchCore, patchCircle, patchCont, patchEnn, patchFDI, patchGyrate, patchNcore, patchPara, patchPerim, patchShape) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=F)
```


Answer Challenge 1 with your written text here.


## Challenge 2 (4 points)

**In our lab, we used the 8-direction or "queen" rule to delineate patches. Using the nlcdSimple raster we created, explore the differences in patch characteristics if you were to use the 4-direction or "rook" rule for delineating patches. Calculate the following class-level metrics for forest cover for both the queen and rook patch delineation rules: number of patches, mean patch size, standard deviation in patch size, nearest-neighbor distance, and total edge length. What patterns do you notice? When might it be appropriate to use one rule vs. the other?**

```{r}
classMnPatch8 = lsm_c_area_mn(forest, directions=8)
classMnPatch8
classMnPatch4 = lsm_c_area_mn(forest, directions=4) 
classMnPatch4
classSdPatch8 = lsm_c_area_sd(forest, directions=8) 
classSdPatch8
classSdPatch4 = lsm_c_area_sd(forest, directions=4) 
classSdPatch4
classTePatch8 = lsm_c_te(forest, count_boundary = TRUE, directions=8) 
classTePatch8
classTePatch4 = lsm_c_te(forest, count_boundary = TRUE, directions=4) 
classTePatch4
classEnnPatch8 = lsm_c_enn_mn(forest, directions=8) 
classEnnPatch8
classEnnPatch4 = lsm_c_enn_mn(forest, directions=4) 
classEnnPatch4
classNumPatch8 = lsm_c_np(forest, directions=8) 
classNumPatch8
classNumPatch4 = lsm_c_np(forest, directions=4) 
classNumPatch4

```

All metrics are markedly higher in the queens case, other than number of patches. I suspect total edge would be higher in the queens case as well, but it keeps showing no change... Using one rule over the other should be based on the context of the situation. This could be based on what the species in question is and it's dispersion abilities, as well as the resolution of the grains to determine what a suitable corridor would be for that species. When working with small grain sizes that should a narrow sub patch between slope wetland patches, it's likely that this is just a densely vegetated ditch with a narrow riparian buffer. I'd expect amphibians to utilize the channels for dispersion, but not a bear.


## Challenge 3 (4 points)


**Using the same zoomed-in study area that we used in the lab, download NLCD raster data for the years 2001 and 2019 and simplify these rasters into 6 landcover categories (wet, developed, barren, forest, grassland, and agriculture). Plot these two rasters. What are some of the initial changes you notice between 2001 and 2019?**

```{r}
studyArea = st_read('/vsicurl/https://github.com/ValenteJJ/SpatialEcology/raw/main/Week4/studyArea.shp')
nlcd2001 = get_nlcd(studyArea, label='AlLandscape', year=2001)

# values(nlcd) = as.character(values(nlcd))

nlcdSimple2001 = nlcd2001
nlcdSimple2001[nlcdSimple2001==11] = 1 #Wet areas are a 1 now
nlcdSimple2001[nlcdSimple2001 %in% c(21, 22, 23, 24)] = 2 #All developed areas are 2
nlcdSimple2001[nlcdSimple2001 %in% c(31, 52)] = 3 #Barren land and shrub/scrub are 3
nlcdSimple2001[nlcdSimple2001 %in% c(41,42,43)] = 4 #All forest types are 4
nlcdSimple2001[nlcdSimple2001 == 71] = 5 #Grassland is 5
nlcdSimple2001[nlcdSimple2001 %in% c(81,82)] = 6 #And agriculture is 6

#Give these numbers category names
tmp = data.frame('ID' = c(1, 2, 3, 4, 5, 6),
                 'category' = c('wetland', 'developed', 'open', 'forest', 'grassland', 'agriculture'))
nlcdSimple2001 = categories(nlcdSimple2001, value=tmp)

#And plot the new raster
ggplot(nlcdSimple2001, aes(x=x, y=y, fill=category)) +
  geom_raster()+
  scale_fill_manual(values=c('blue', 'black', 'gray', 'darkolivegreen', 'orange', 'yellow'))

nlcd2019 = get_nlcd(studyArea, label='AlLandscape', year=2019)

# values(nlcd) = as.character(values(nlcd))

nlcdSimple2019 = nlcd2019
nlcdSimple2019[nlcdSimple2019==11] = 1 #Wet areas are a 1 now
nlcdSimple2019[nlcdSimple2019 %in% c(21, 22, 23, 24)] = 2 #All developed areas are 2
nlcdSimple2019[nlcdSimple2019 %in% c(31, 52)] = 3 #Barren land and shrub/scrub are 3
nlcdSimple2019[nlcdSimple2019 %in% c(41,42,43)] = 4 #All forest types are 4
nlcdSimple2019[nlcdSimple2019 == 71] = 5 #Grassland is 5
nlcdSimple2019[nlcdSimple2019 %in% c(81,82)] = 6 #And agriculture is 6

#Give these numbers category names
tmp = data.frame('ID' = c(1, 2, 3, 4, 5, 6),
                 'category' = c('wetland', 'developed', 'open', 'forest', 'grassland', 'agriculture'))
nlcdSimple2019 = categories(nlcdSimple2019, value=tmp)

#And plot the new raster
ggplot(nlcdSimple2019, aes(x=x, y=y, fill=category)) +
  geom_raster()+
  scale_fill_manual(values=c('blue', 'black', 'gray', 'darkolivegreen', 'orange', 'yellow'))


res(nlcdSimple2001)
ext(nlcdSimple2001)
levels(nlcdSimple2001)
res(nlcdSimple2019)
ext(nlcdSimple2019)
levels(nlcdSimple2019)

```

The main trend is agriculture expanding into previously forested land. I suspect that just like how the power line trail of early succession habitat was classified as agriculture rather than grassland, the same is likely true of the grassland patches and these are actually ag as well. There's a lot more barren land around these grassland patches though, so this could be recently harvest forests where development is occurring around in phases. 

**Quantify this at the class level by calculating and reporting the changes in (1) the total amount of each land cover type (2) mean patch size for each land cover type, and (3) mean nearest neighbor distance for each cover type between the years 2011 and 2019. Give a short description of how you interpret the changes in these values.**

```{r}
classMnPatch2001 = lsm_c_area_mn(nlcdSimple2001, directions=8)  

classMnPatch2001

classMnPatch2019 = lsm_c_area_mn(nlcdSimple2019, directions=8) 

classMnPatch2019

classEnnPatch2001 = lsm_c_enn_mn(nlcdSimple2001, directions=8) 

classEnnPatch2001

classEnnPatch2019 = lsm_c_enn_mn(nlcdSimple2019, directions=8) 

classEnnPatch2019

classArea2001 = lsm_c_ca(nlcdSimple2001, directions = 8) 

classArea2001

classArea2019 = lsm_c_ca(nlcdSimple2019, directions = 8) 

classArea2019
```

mean patch size area 2001-2019:
Wetland = 0, developed = 0.18, open = 2.53363, forest = -31.16128, grassland = 1.766824, ag = 4.780793 
mean nearest neighbor:
wetland = 0, developed = -6.41816, open = -197.8003, forest = -4.26063, grassland = -64.44042, ag = -3.11571
total area:
wetland = 0, developed = 1.35, open = 159.39, forest = -413.46, grassland = 125.64, ag = 127.08


**Quantify these changes at the landscape level by calculating and reporting on changes in the (1) Shannon diversity and (2) Shannon evenness of the landscapes at the different time points. Give a short description of how you interpret the changes in these values.**

```{r}
ShanDI2001 = lsm_l_shdi(nlcdSimple2001)
ShanDI2019 = lsm_l_shdi(nlcdSimple2019)
ShanDI2001
ShanDI2019

ShanEI2001 = lsm_l_shei(nlcdSimple2001)
ShanEI2019 = lsm_l_shei(nlcdSimple2019)
ShanEI2001
ShanEI2019
```

Diversity Index values increased between 2001-2019, indicating the diversity was lowered. No land cover types went away completely, so I think this was influenced by a change in abundance for each cover type. Evenness values also increased in these years from 0.53 to 0.75, which I believe means that areas of different groups became more evenly represented so that total area on a class basis is closer to an even distribution now.  



## Challenge 4 (4 points)

**Use the voss2d() function to simulate a surface where g = 7 and H = 0.5. From that, create 9 'landscapes' with 10%, 20%, 30%, ..., 90% threshold values. The '1' values here can represent anything your imagination comes up with. It could be forest cover, cover of some other land cover type, bodies of water, temperatures above a threshold, etc. I suggest you set the seed for your simulation value so that you get the same outcome each time you run the code. Plot these landscapes and comment on what patterns you see changing as the value increases from a 10% cover to 90% cover.**

```{r}
set.seed(23)

vossModel = voss2d(g=7, H=0.5)
vossModel = rast(vossModel$z)

plot(vossModel)

# 10% cover
threshold10 = quantile(as.matrix(vossModel), prob=0.1)
voss10 = ifel(vossModel > threshold10, 0, 1)
plot(voss10)

# 20% cover
threshold20 = quantile(as.matrix(vossModel), prob=0.2)
voss20 = ifel(vossModel > threshold20, 0, 1)
plot(voss20)

# 30% cover
threshold30 = quantile(as.matrix(vossModel), prob=0.3)
voss30 = ifel(vossModel > threshold30, 0, 1)
plot(voss30)

# 40% cover
threshold40 = quantile(as.matrix(vossModel), prob=0.4)
voss40 = ifel(vossModel > threshold40, 0, 1)
plot(voss40)

# 50% cover
threshold50 = quantile(as.matrix(vossModel), prob=0.5)
voss50 = ifel(vossModel > threshold50, 0, 1)
plot(voss50)

# 60% cover
threshold60 = quantile(as.matrix(vossModel), prob=0.6)
voss60 = ifel(vossModel > threshold60, 0, 1)
plot(voss60)

# 70% cover
threshold70 = quantile(as.matrix(vossModel), prob=0.7)
voss70 = ifel(vossModel > threshold70, 0, 1)
plot(voss70)

# 80% cover
threshold80 = quantile(as.matrix(vossModel), prob=0.8)
voss80 = ifel(vossModel > threshold80, 0, 1)
plot(voss80)

# 90% cover
threshold90 = quantile(as.matrix(vossModel), prob=0.9)
voss90 = ifel(vossModel > threshold90, 0, 1)
plot(voss90)
```

The pattern I see is that aggregation is happening near the edges and fills into the center last. Where converging masses of 'land cover' occur, each threshold increase fills into the spaces between them and convergence just becomes more of a thing. It's not like new values are popping up in the center away from everything.

**Identify 3 class-level or landscape-level metrics that help you capture the changes you are observing. Calculate those metrics for each of the 9 landscapes and plot them (x-axis is threshold value and y-axis is calculated metric). Briefly describe why you chose these 3 metrics and how they change with increasing cover.**

```{r}

vossTe10 = lsm_c_te(voss10, count_boundary = TRUE, directions=8) 
vossTe10
vossTe20 = lsm_c_te(voss20, count_boundary = TRUE, directions=8) 
vossTe20
vossTe30 = lsm_c_te(voss30, count_boundary = TRUE, directions=8) 
vossTe30
vossTe40 = lsm_c_te(voss40, count_boundary = TRUE, directions=8) 
vossTe40
vossTe50 = lsm_c_te(voss50, count_boundary = TRUE, directions=8) 
vossTe50
vossTe60 = lsm_c_te(voss60, count_boundary = TRUE, directions=8) 
vossTe60
vossTe70 = lsm_c_te(voss70, count_boundary = TRUE, directions=8) 
vossTe70
vossTe80 = lsm_c_te(voss80, count_boundary = TRUE, directions=8) 
vossTe80
vossTe90 = lsm_c_te(voss90, count_boundary = TRUE, directions=8) 
vossTe90
yTe <-c(1706, 2056, 2796, 3554, 3644, 3254, 2938, 2550, 1966)
xThresh <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Te_vs_Thresh = plot(x= xThresh, y=yTe)
Te_vs_Thresh

vossEnn_mn_10 = lsm_c_enn_mn(voss10, directions=8)
vossEnn_mn_10
vossEnn_mn_20 = lsm_c_enn_mn(voss20, directions=8)
vossEnn_mn_20
vossEnn_mn_30 = lsm_c_enn_mn(voss30, directions=8)
vossEnn_mn_30
vossEnn_mn_40 = lsm_c_enn_mn(voss40, directions=8)
vossEnn_mn_40
vossEnn_mn_50 = lsm_c_enn_mn(voss50, directions=8)
vossEnn_mn_50
vossEnn_mn_60 = lsm_c_enn_mn(voss60, directions=8)
vossEnn_mn_60
vossEnn_mn_70 = lsm_c_enn_mn(voss70, directions=8)
vossEnn_mn_70
vossEnn_mn_80 = lsm_c_enn_mn(voss80, directions=8)
vossEnn_mn_80
vossEnn_mn_90 = lsm_c_enn_mn(voss90, directions=8)
vossEnn_mn_90
yEnn_mn <-c(3.199507, 3.637775, 2.939403, 2.540016, 2.535572, 2.687083, 2.663815, 2.413240, 2.244642)
xThresh <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Enn_mn_vs_Thresh = plot(x= xThresh, y=yEnn_mn)
Enn_mn_vs_Thresh

voss_ED_10 = lsm_l_ed(voss10, directions = 8)
voss_ED_10
voss_ED_20 = lsm_l_ed(voss20, directions = 8)
voss_ED_20
voss_ED_30 = lsm_l_ed(voss30, directions = 8)
voss_ED_30
voss_ED_40 = lsm_l_ed(voss40, directions = 8)
voss_ED_40
voss_ED_50 = lsm_l_ed(voss50, directions = 8)
voss_ED_50
voss_ED_60 = lsm_l_ed(voss60, directions = 8)
voss_ED_60
voss_ED_70 = lsm_l_ed(voss70, directions = 8)
voss_ED_70
voss_ED_80 = lsm_l_ed(voss80, directions = 8)
voss_ED_80
voss_ED_90 = lsm_l_ed(voss90, directions = 8)
voss_ED_90
yED <-c(953.6686, 1127.937, 1540.773, 1961.421, 1990.866, 1735.473, 1524.548, 1269.755, 900.7872)
xThresh <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Enn_ED_vs_Thresh = plot(x= xThresh, y=yED)
Enn_ED_vs_Thresh
```

I first chose Total edge, because I was curious what the pattern would look like exactly. I expected it to go up and down because core would over take edge as cores grew. Normal distribution, No surprise here. The pattern I observed for distance to nearest neighbor was a little more oscillating between small changes in x, but overall had a negative relationship with increasing thresholds. This one I chose because I expected the negative relationship as patches got bigger, but really expected something more like a logistic looking regression. From the landscape level, I chose edge density for the same reasons as total edge, but for the sake of seeing how area would influence this metric. It appears to have an almost normal distribution when plotted with threshold, though there's more of a quick rise and slow fall in terms of effect, so sort of right skewed. This adds up, since more area means more edge before patches converge on the rising limb, and area values get larger and edge values get smaller after convergence, on the falling limb.

## Challenge 5 (4 points)

**Use the voss2d() function to simulate 9 surfaces. In each one, g should be 7, but vary the value of H from 0.1 to 0.9. Create a unique landscape from each of these with a threshold value of 30% cover. Again, the 1 values can represent anything you desire, and again I suggest you set the seed for your simulation value so that you get the same outcome each time you run the code. Plot these landscapes and comment on what patterns you see changing as the H value increases from 0.1 to 0.9.**

```{r}
set.seed(23)

vossModelH01 = voss2d(g=7, H=0.1)
vossModelH01 = rast(vossModelH01$z)
vossModelH02 = voss2d(g=7, H=0.2)
vossModelH02 = rast(vossModelH02$z)
vossModelH03 = voss2d(g=7, H=0.3)
vossModelH03 = rast(vossModelH03$z)
vossModelH04 = voss2d(g=7, H=0.4)
vossModelH04 = rast(vossModelH04$z)
vossModelH05 = voss2d(g=7, H=0.5)
vossModelH05 = rast(vossModelH05$z)
vossModelH06 = voss2d(g=7, H=0.6)
vossModelH06 = rast(vossModelH06$z)
vossModelH07 = voss2d(g=7, H=0.7)
vossModelH07 = rast(vossModelH07$z)
vossModelH08 = voss2d(g=7, H=0.8)
vossModelH08 = rast(vossModelH08$z)
vossModelH09 = voss2d(g=7, H=0.9)
vossModelH09 = rast(vossModelH09$z)

plot(vossModelH01)
plot(vossModelH02)
plot(vossModelH03)
plot(vossModelH04)
plot(vossModelH05)
plot(vossModelH06)
plot(vossModelH07)
plot(vossModelH08)
plot(vossModelH09)


H01threshold30 = quantile(as.matrix(vossModelH01), prob=0.3)
H01 = ifel(vossModelH01 > H01threshold30, 0, 1)
plot(H01)

H02threshold30 = quantile(as.matrix(vossModelH02), prob=0.3)
H02 = ifel(vossModelH02 > H02threshold30, 0, 1)
plot(H02)

H03threshold30 = quantile(as.matrix(vossModelH03), prob=0.3)
H03 = ifel(vossModelH03 > H03threshold30, 0, 1)
plot(H03)

H04threshold30 = quantile(as.matrix(vossModelH04), prob=0.3)
H04 = ifel(vossModelH04 > H04threshold30, 0, 1)
plot(H04)

H05threshold30 = quantile(as.matrix(vossModelH05), prob=0.3)
H05 = ifel(vossModelH05 > H05threshold30, 0, 1)
plot(H05)

H06threshold30 = quantile(as.matrix(vossModelH06), prob=0.3)
H06 = ifel(vossModelH06 > H06threshold30, 0, 1)
plot(H06)

H07threshold30 = quantile(as.matrix(vossModelH07), prob=0.3)
H07 = ifel(vossModelH07 > H07threshold30, 0, 1)
plot(H07)

H08threshold30 = quantile(as.matrix(vossModelH08), prob=0.3)
H08 = ifel(vossModelH08 > H08threshold30, 0, 1)
plot(H08)

H09threshold30 = quantile(as.matrix(vossModelH09), prob=0.3)
H09 = ifel(vossModelH09 > H09threshold30, 0, 1)
plot(H09)

```

Basically, land cover just becomes more spatially aggregated as H increases.

**Identify 3 class-level or landscape-level metrics that help you capture the changes you are observing. THESE MUST BE DIFFERENT THAN THOSE METRICS YOU USED IN CHALLENGE 2. Calculate those metrics for each of the 9 landscapes and plot them (x-axis is H-value and y-axis is calculated metric). Briefly describe why you chose these 3 metrics and how they change with increasing cover.**

```{r}

H01_np = lsm_l_np(H01, directions=8) 
H01_np
H02_np = lsm_l_np(H02, directions=8) 
H02_np
H03_np = lsm_l_np(H03, directions=8) 
H03_np
H04_np = lsm_l_np(H04, directions=8) 
H04_np
H05_np = lsm_l_np(H05, directions=8) 
H05_np
H06_np = lsm_l_np(H06, directions=8) 
H06_np
H07_np = lsm_l_np(H07, directions=8) 
H07_np
H08_np = lsm_l_np(H08, directions=8) 
H08_np
H09_np = lsm_l_np(H09, directions=8) 
H09_np
yNP <-c(370, 262, 147, 186, 150, 123, 38, 31, 25)
xH <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
NP_vs_H = plot(x= xH, y=yNP)
NP_vs_H

H01_circ = lsm_l_circle_mn(H01, directions = 8)
H01_circ
H02_circ = lsm_l_circle_mn(H02, directions = 8)
H02_circ
H03_circ = lsm_l_circle_mn(H03, directions = 8)
H03_circ
H04_circ = lsm_l_circle_mn(H04, directions = 8)
H04_circ
H05_circ = lsm_l_circle_mn(H05, directions = 8)
H05_circ
H06_circ = lsm_l_circle_mn(H06, directions = 8)
H06_circ
H07_circ = lsm_l_circle_mn(H07, directions = 8)
H07_circ
H08_circ = lsm_l_circle_mn(H08, directions = 8)
H08_circ
H09_circ = lsm_l_circle_mn(H09, directions = 8)
H09_circ
yCirc <-c(0.4842729, 0.4720764, 0.4661469, 0.4813961, 0.4650786, 0.4581993, 0.46638, 0.4738502, 0.5126833)
xH <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Circ_vs_H = plot(x= xH, y=yCirc)
Circ_vs_H

H01_frac = lsm_c_frac_mn(H01, directions = 8)
H01_frac
H02_frac = lsm_c_frac_mn(H02, directions = 8)
H02_frac
H03_frac = lsm_c_frac_mn(H03, directions = 8)
H03_frac
H04_frac = lsm_c_frac_mn(H04, directions = 8)
H04_frac
H05_frac = lsm_c_frac_mn(H05, directions = 8)
H05_frac
H06_frac = lsm_c_frac_mn(H06, directions = 8)
H06_frac
H07_frac = lsm_c_frac_mn(H07, directions = 8)
H07_frac
H08_frac = lsm_c_frac_mn(H08, directions = 8)
H08_frac
H09_frac = lsm_c_frac_mn(H09, directions = 8)
H09_frac
yfrac <-c(1.230115, 1.230203, 1.183577, 1.198105, 1.198105, 1.163642, 1.081681, 1.104732, 1.256054)
xH <-c(10, 20, 30, 40, 50, 60, 70, 80, 90)
frac_vs_H = plot(x= xH, y=yfrac)
frac_vs_H
```

Number of patches had a similar pattern to nearest neighbor distance in the last example, which was a little surprising. I was expecting something more linear, and other than chance, I'm not sure why this is the case. Related circumscribing circle (mean) surprised me, as it didn't really have any kind of relationship with H. With area being consistent between all models, I thought this one would capture changes in shape pretty well. Something else going on that I'm not picking up on, I think. Tried this with mean fractional dimension index and got similar results as well. I had similar expectations as those I had with circumscribing circle where I thought with area controlled for, shape would have some linear relationship with H which seems to aggregate land cover as values rise. I'm not sure what I'm missing here or if I'm overlooking some part of the math each function uses.

